---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  word_document: default
---


# Loading and preprocessing the data

#### Read data
```{r, echo = TRUE}
setwd("D:/Course5_wk2_assignment")
activity_data <- read.csv("./activity.csv")
str(activity_data)
```

#### Transform date from Factor to Date class 
```{r, echo = TRUE}
activity_data$date <- as.Date(activity_data$date, format = "%Y-%m-%d")
```

# What is mean total number of steps taken per day?

#### Calculate steps per day
```{r, echo = TRUE}
steps_per_day  <- tapply(activity_data$steps, activity_data$date, sum, na.rm= TRUE)
```

#### Make a histogram of the total number of steps taken each day
```{r, echo = TRUE}
hist(steps_per_day, col = "red", breaks=20, xlab = "total steps per day", main ="Histogram total number of steps per day" )
```

#### Calculate and report mean and median of the total number of steps taken per day
```{r, echo = TRUE}
mean_per_day  <- round(mean(steps_per_day, na.rm = TRUE))
median_per_day  <- round(median(steps_per_day, na.rm = TRUE))

print(mean_per_day)

print(median_per_day)

```

# What is the average daily activity pattern?

#### Calculate average steps
```{r,echo = TRUE}
average_steps <- aggregate(activity_data$steps, list(interval=activity_data$interval), mean, na.rm = TRUE)
```
 
#### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)  

```{r,echo = TRUE}     
plot(average_steps, type = "l", col = "red", main = "average number of steps of the 5-minute interval", xlab = "5-minute interval", ylab = "average number of steps")
```

#### Determine the 5-minute interval with the maximum number of steps
```{r,echo = TRUE}
max_steps <- max(average_steps$x)
max_interval <- average_steps[which(average_steps$x == max_steps),1]

print(max_interval)

```

# Imputing missing values

#### Calculate and report the total number of missing values in the dataset
```{r, echo = TRUE}
total_NA <- sum(is.na(activity_data))

print(total_NA)
```

#### Create a new dataset which is a copy of the original dataset
```{r, echo = TRUE}
activity_data_nw <- activity_data
```

#### Replace NA with the average of the interval.
```{r, echo = TRUE}
activity_data_nw$steps[which(is.na(activity_data_nw$steps))] = rep(average_steps$x, 288)[which(is.na(activity_data_nw$steps))]
```

####  Calculate total number of steps per day fot the new data
```{r, echo = TRUE}
steps_per_day_nw  <- tapply(activity_data_nw$steps, activity_data_nw$date, sum, na.rm= TRUE)
```

#### Make a histogram of the total number of steps taken each day of the new data
```{r, echo = TRUE}
hist(steps_per_day_nw, col = "red", breaks=20, xlab = "total steps per day", main ="Histogram of total number of steps per day (NA replaced by mean)" )
```

#### Calculate the mean and median of the new data
```{r, echo = TRUE}
mean_per_day_nw  <- mean(steps_per_day_nw, na.rm = TRUE)
median_per_day_nw  <- median(steps_per_day_nw, na.rm = TRUE)

print(mean_per_day_nw)

print(median_per_day_nw)
```

#### Calculate the differences between mean and median based on the new and old data
```{r, echo = TRUE}
diff_mean <- mean_per_day_nw - mean_per_day
diff_median <- median_per_day_nw - median_per_day

print(diff_mean)

print(diff_median)
```

#### The mean and median based on the data, where NAs are replaced by mean are larger than the mean and median based on original data.


# Are there differences in activity patterns between weekdays and weekends?

#### define weekdays 
```{r, echo= TRUE}
## setlocale
Sys.setlocale("LC_TIME", "English")

days <- weekdays(activity_data_nw$date)
activity_data_days <- cbind(activity_data_nw, days)

week_day <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
activity_data_days$dayType = ifelse(days %in% week_day, "weekday", "weekend")
```

#### Calculate average steps for weekdays and weekends
```{r, echo= TRUE}
average_steps <- aggregate(activity_data_days$steps, list(interval=activity_data_days$interval, dayType = activity_data_days$dayType), mean, na.rm = TRUE)
```

#### Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days
```{r, echo= TRUE}
library(ggplot2)
g <- ggplot(average_steps, aes(interval, x))
g <- g + facet_grid(dayType~.)
p <- g + geom_line(stat="identity", color = "red")+labs(x="interval",y="number of steps")+labs(title="activity patterns between weekdays and weekends")
print(p)
```

#### There are differences between activity pattern weekdays and weekends.


