---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


# Loading and preprocessing the data

```{r, echo = TRUE}
# read data
setwd("D:/Course5_wk2_assignment")
activity_data <- read.csv("./activity.csv")
str(activity_data)

# Transform date from Factor to Date class 
activity_data$date <- as.Date(activity_data$date, format = "%Y-%m-%d")
```

# What is mean total number of steps taken per day?
```{r, echo = TRUE}
steps_per_day  <- tapply(activity_data$steps, activity_data$date, sum, na.rm= TRUE)

# make histogram
hist(steps_per_day, col = "red", breaks=20, xlab = "total steps per day", main ="Histogram total number of steps per day" )

# calculate and report mean and median of the total number of steps taken per day
mean_per_day  <- round(mean(steps_per_day, na.rm = TRUE))
median_per_day  <- round(median(steps_per_day, na.rm = TRUE))

# mean of total number of steps per day
print(mean_per_day)

# median of total number of steps per day
print(median_per_day)

```

# What is the average daily activity pattern?
```{r}
# calculate average steps
average_steps <- aggregate(activity_data$steps, list(interval=activity_data$interval), mean, na.rm = TRUE)
     
# make plot
plot(average_steps, type = "l", col = "red", main = "average number of steps of the 5-minute interval", xlab = "5-minute interval", ylab = "average number of steps")

# the 5-minute interval with the maximum number of steps
max_steps <- max(average_steps$x)
max_interval <- average_steps[which(average_steps$x == max_steps),1]

# interval with maximum number of steps
print(max_interval)

```

# Imputing missing values
```{r}
# Calculate and report the total number of missing values in the dataset 
total_NA <- sum(is.na(activity_data))

# number of missing values in the dataset
print(total_NA)

# create a new dataset which is a copy of the original dataset
activity_data_nw <- activity_data

# replace NA with the average of the interval.
activity_data_nw$steps[which(is.na(activity_data_nw$steps))] = rep(average_steps$x, 288)[which(is.na(activity_data_nw$steps))]

# create a new dataset which is a copy of the original dataset
activity_data_nw <- activity_data

# replace NA with the average of the interval.
activity_data_nw$steps[which(is.na(activity_data_nw$steps))] = rep(average_steps$x, 288)[which(is.na(activity_data_nw$steps))]

# calculate total number of steps per day fot the new data
steps_per_day_nw  <- tapply(activity_data_nw$steps, activity_data_nw$date, sum, na.rm= TRUE)

# make a histogram of the new data
hist(steps_per_day_nw, col = "red", breaks=20, xlab = "total steps per day", main ="Histogram of total number of steps per day (NA replaced by mean)" )

# calculate the mean and median of the new data
mean_per_day_nw  <- mean(steps_per_day_nw, na.rm = TRUE)
median_per_day_nw  <- median(steps_per_day_nw, na.rm = TRUE)

# mean of total number of steps per day of the data NA replaced by mean
print(mean_per_day_nw)

# median of total number of steps per day of the data NA replaced by mean
print(median_per_day_nw)

# calculate the differences
diff_mean <- mean_per_day_nw - mean_per_day
diff_median <- median_per_day_nw - median_per_day

# difference between the both means
print(diff_mean)

# difference between the both medians
print(diff_median)
```

#### The mean and median based on the data, where NAs are replaced by mean are larger than the mean and median based on original data.


#Are there differences in activity patterns between weekdays and weekends?

```{r}
## setlocale
Sys.setlocale("LC_TIME", "English")

days <- weekdays(activity_data_nw$date)
activity_data_days <- cbind(activity_data_nw, days)

# define weekdays    
week_day <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
activity_data_days$dayType = ifelse(days %in% week_day, "weekday", "weekend")

# calculate average steps for weekdays and weekends
average_steps <- aggregate(activity_data_days$steps, list(interval=activity_data_days$interval, dayType = activity_data_days$dayType), mean, na.rm = TRUE)

# make plot
library(ggplot2)
g <- ggplot(average_steps, aes(interval, x))
g <- g + facet_grid(dayType~.)
p <- g + geom_line(stat="identity", color = "red")+labs(x="interval",y="number of steps")+labs(title="activity patterns between weekdays and weekends")
print(p)
```

#### There are differences between activity pattern weekdays and weekends.


